An algorithm for finding the global extremum\\ of a partially defined function\thanks{This work was supported by the Russian Science Foundation, project No. 21-11-00204.}

### 2023-10-06 2023/MMST/samplepaper.tex
add one more reference

However, a principally new problem in applications, namely the numerical instability of the models investigated in some (unknown a priori) subregions of the parameters' variations recently became relevant. In such subregions, one cannot perform the numerical modeling and compute the objective function values correctly. This phenomenon can be interpreted either as the presence of some hidden constraints \cite{Stripinis2021} or as a partial computability of the objective function in the search domain \cite{Candelieri2019,Sergeyev2003,Strongin2020}. In such a statement, the optimization problem becomes more complex since the region of feasible parameters' combinations is not defined in advance.

@@

\subsubsection{The third approach.} To restore the objective function values at the points of the intervals with two non-computable points as the averages of the neighbor points and to apply corresponding computation rule from formula (\ref{eq9}) assuming the points to be computable ones. If the neighbor points are non-computable as well, use formula (\ref{eq14}). In the case of an interval with one non-computable point, apply formulae (\ref{eq15}) and (\ref{eq16}).
\begin{figure}
\includegraphics[width=\textwidth]{fig3.png}
\caption{Imputation of the non-computable values by the averages of the neighbor points. Approach No. 3} \label{fig3}
\end{figure}

@@

\subsubsection{The fourth approach.} To restore the objective function values at the points of the intervals with two non-computable points as the averages of the nearest computable points taking into account the coordinates, i.e.
\begin{equation}\label{eq17} 
%\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \| x-x' \|}{\| x''-x' \|},
\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \rho(x,x')}{\rho(x',x'') },
\end{equation}
where $z'$ and $z''$ are the objective function values in the computable points $x'$ and $x''$  nearest to the interval $(x_{i-1},x_i)$ and $\rho(x_1,x_2) =  |x_1 - x_2|^{1/N}$.
The rest of the approach is identical to the third one. Formula (\ref{eq14}) in this case will be applied when the non-computable region is located only near the boundary point.
\begin{figure}
\includegraphics[width=\textwidth]{fig4.png}
\caption{Imputation of non-computable values by averages of the neighboring points. Approach No. 4} \label{fig4}
\end{figure}
\begin{figure}
\includegraphics[width=\textwidth]{fig5.png}
\caption{Imputation of non-computable values by averages of the minorant and the majorant. Approach No. 5} \label{fig5}
\end{figure}

>>>>

However, a principally new problem in applications, namely the numerical instability of the models investigated in some (unknown a priori) subregions of the parameters' variations recently became relevant. In such subregions, one cannot perform the numerical modeling and compute the objective function values correctly. This phenomenon can be interpreted either as the presence of some hidden constraints \cite{Stripinis2021}, or as unknown discontinuous regions \cite{Audet2022}, or as a partial computability of the objective function in the search domain \cite{Candelieri2019,Sergeyev2003,Strongin2020}. In such a statement, the optimization problem becomes more complex since the region of feasible parameters' combinations is not defined in advance.

@@

\subsubsection{The third approach.} To restore the objective function values at the points of the intervals with two non-computable points as the averages of the neighbor points and to apply corresponding computation rule from formula (\ref{eq9}) assuming the points to be computable ones. If the neighbor points are non-computable as well, use formula (\ref{eq14}). In the case of an interval with one non-computable point, apply formulae (\ref{eq15}) and (\ref{eq16}).
\begin{figure}
\centering
\includegraphics[width=0.95\textwidth]{fig3.png}
\caption{Imputation of the non-computable values by the averages of the neighbor points. Approach No. 3} \label{fig3}
\end{figure}

@@

\subsubsection{The fourth approach.} To restore the objective function values at the points of the intervals with two non-computable points as the averages of the nearest computable points taking into account the coordinates, i.e.
\begin{equation}\label{eq17} 
%\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \| x-x' \|}{\| x''-x' \|},
\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \rho(x,x')}{\rho(x',x'') },
\end{equation}
where $z'$ and $z''$ are the objective function values in the computable points $x'$ and $x''$  nearest to the interval $(x_{i-1},x_i)$ and $\rho(x_1,x_2) =  |x_1 - x_2|^{1/N}$.
The rest of the approach is identical to the third one. Formula (\ref{eq14}) in this case will be applied when the non-computable region is located only near the boundary point.
\begin{figure}
\centering
\includegraphics[width=0.95\textwidth]{fig4.png}
\caption{Imputation of non-computable values by averages of the neighboring points. Approach No. 4} \label{fig4}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.95\textwidth]{fig5.png}
\caption{Imputation of non-computable values by averages of the minorant and the majorant. Approach No. 5} \label{fig5}
\end{figure}
